quollr: An R Package for Visualizing 2-D Models from Non-linear Dimension Reductions in High Dimensional Space

### 2025-05-12 paper-quollr.Rmd
Updated the introduction

This paper presents the R package, `quollr` which introduce a new visual tool in determining which Non-linear dimension reduction (NLDR) technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data. The methodology of the algorithm is explained in *cite the methodology paper*. Furthermore, the `quollr` package enables users to perform hexagonal binning [@dan2023], resulting in the generation of regular hexagons. The software is available from the Comprehensive R Archive Network (CARN) at [https://CRAN.R-project.org/package=quollr](https://CRAN.R-project.org/package=quollr).

Add the image generated with the application

@@

```{r}
nldr1_org + nldr2_org + nldr3_org + 
  nldr4_org + nldr5_org + nldr6_org +
  plot_layout(ncol = 3)
```

>>>>

Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) [@laurens2008], uniform manifold approximation and projection (UMAP) [@leland2018], potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm [@moon2019], large-scale dimensionality reduction Using triplets (TriMAP) [@amid2019], and pairwise controlled manifold approximation (PaCMAP) [@yingfan2021], create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \@ref(fig:example)). It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.

This paper presents the R package, `quollr` which introduce a new visual tool to evaluate which NLDR method and which (hyper)parameter choice gives the most accurate representation of high-dimensional data. The methodology of this algorithm is explained in *cite the methodology paper*. The software is available from the Comprehensive R Archive Network (CARN) at [https://CRAN.R-project.org/package=quollr](https://CRAN.R-project.org/package=quollr).

@@

```{r fig-example, fig.alt = "NLDR representations with different methods and (hyper)parameter choices.", fig.cap = "", out.width = "100%", fig.width= 6, fig.height= 4}

nldr1_org + nldr2_org + nldr3_org + 
  nldr4_org + nldr5_org + nldr6_org +
  plot_layout(ncol = 3)
```

### 2025-05-12 paper-quollr.tex
Updated the introduction

This paper presents the R package, \texttt{quollr} which introduce a new visual tool in determining which Non-linear dimension reduction (NLDR) technique and which (hyper)parameter choice gives most accurate representation of high-dimensional data. The methodology of the algorithm is explained in \emph{cite the methodology paper}. Furthermore, the \texttt{quollr} package enables users to perform hexagonal binning \citep{dan2023}, resulting in the generation of regular hexagons. The software is available from the Comprehensive R Archive Network (CARN) at \url{https://CRAN.R-project.org/package=quollr}.

Add the image generated with the application

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-1-1}

@@

@@

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-28-1}

>>>>

Non-linear dimension reduction (NLDR) techniques, such as t-distributed stochastic neighbor embedding (tSNE) \citep{laurens2008}, uniform manifold approximation and projection (UMAP) \citep{leland2018}, potential of heat-diffusion for affinity-based trajectory embedding (PHATE) algorithm \citep{moon2019}, large-scale dimensionality reduction Using triplets (TriMAP) \citep{amid2019}, and pairwise controlled manifold approximation (PaCMAP) \citep{yingfan2021}, create wildly different representations depending on the selected method and (hyper)parameter choices (Figure \ref{fig:example}). It is difficult to determine whether any of these representations are accurate, which one is the best, or whether they have missed important structures.

This paper presents the R package, \texttt{quollr} which introduce a new visual tool to evaluate which NLDR method and which (hyper)parameter choice gives the most accurate representation of high-dimensional data. The methodology of this algorithm is explained in \emph{cite the methodology paper}. The software is available from the Comprehensive R Archive Network (CARN) at \url{https://CRAN.R-project.org/package=quollr}.

\begin{figure}
\includegraphics[width=1\linewidth,alt={NLDR representations with different methods and (hyper)parameter choices.}]{paper-quollr_files/figure-latex/fig-example-1} \end{figure}

@@

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-18-1}

@@

\includegraphics[width=1\linewidth]{paper-quollr_files/figure-latex/unnamed-chunk-27-1}
